\titleNameLong{}

### 2015-12-08 FinalProject/Tex/paper/paper.tex
added clarifying wording in introduction
removed contractions

The term GPU was popularized around the turn of the 21st century by NVIDIA as
consumer-level devices began to need dedicated accelerators to render 3D
graphics. Today, most all computers contain GPUs, from supercomputers to
laptops, desktops to smartphones, although some of these do not use dedicated
graphics cards. The area of GPGPU started to take off roughly a decade ago when
it began to be practical to solve traditional problems with the massively
parallel nature of GPUs. Using hundreds or even thousands of ``cores" to perform
computations in parallel is tempting (even if the cores aren't really cores at
all), but brings with it a new set of challenges.

In a way, the problems of GPGPU are an extension of those sometimes seen in a
traditional desktop multiprocessing system. Even when under a heavy compute
load, it isn't always easy to keep all of the cores doing useful work in any
given cycle. Between limitations in structural units, dependences between
instructions, and latencies of memory accesses, the processor cannot help but
stall sometimes. This problem is exacerbated by orders of magnitude when solving
heterogeneous problems on a GPU. GPUs were originally designed to operate on
graphics processing, problems that are well-structured and have simple, inherent
data parallelism. Problems that do not contain parallelism that is easily mapped
to the architecture of a GPU fail to keep all of the cores in a GPU busy or
strain the memory bandwidth past the GPU's ability to hide latency.

The remainder of this survey is organized as follows. Section
\ref{sec:architecture} provides a brief architectural overview of GPUs. Section
\ref{sec:challenges} elaborates on the challenges specific to GPGPU
applications. Section \ref{sec:dynamic} discusses spawning granular helpers to
increase performance. Section \ref{sec:scheduling} deals with more efficient
scheduling within and among streaming multiprocessors. Section \ref{sec:memory}
talks specifically about reducing memory contention. Section \ref{sec:software}
is a discussion of a few software and compiler solutions to the challenges of
GPGPU. Section
\ref{sec:conclusions} concludes the survey.

@@

Individual threads exist within a thread-block. These threads are grouped into
``warps'', usually around 32 at a time. The warp is the fundamental unit of
execution on GPUs. Threads within a warp are mapped to a vector functional unit
such that all threads execute the same instruction on different data, just a
like a SIMD unit. With this in mind, it would be more accurate to simply call
``threads'' a sequence of SIMD lane instructions, but this survey will continue
to refer to them as threads to maintain consistency with the literature. The
term ``Warp'' is the term used by NVIDIA, while AMD and OpenCL use
``wavefront''. Maxwell supports up to 64 warps resident per SM. Note that this
doesn't mean that all 64 warps are executing at the same time, but that the
architecture can support that many warp contexts concurrently. The SM maintains
a warp pool to store the context of all running threads. Warps are scheduled by
the warp schedulers from this pool.

>>>>

The term GPU was popularized around the turn of the 21st century by NVIDIA as
consumer-level devices began to need dedicated accelerators to render 3D
graphics. Today, most all computers contain GPUs, from supercomputers to
laptops, desktops to smartphones, although some of these do not use dedicated
graphics cards. The area of GPGPU started to take off roughly a decade ago when
it began to be practical to solve traditional problems with the massively
parallel nature of GPUs. Using hundreds or even thousands of ``cores" to perform
computations in parallel is tempting (even if the cores are not really cores at
all), but brings with it a new set of challenges.

In a way, the problems of GPGPU are an extension of those sometimes seen in a
traditional desktop multiprocessing system. Even when under a heavy compute
load, it is not always easy to keep all of the cores doing useful work in any
given cycle. Between limitations in structural units, dependences between
instructions, and latencies of memory accesses, the processor cannot help but
stall sometimes. This problem is exacerbated by orders of magnitude when solving
heterogeneous problems on a GPU. GPUs were originally designed to operate on
graphics processing, problems that are well-structured and have simple, inherent
data parallelism. Problems that do not contain parallelism that is easily mapped
to the architecture of a GPU fail to keep all of the cores in a GPU busy or
strain the memory bandwidth past the GPU's ability to hide latency.

The remainder of this survey is organized as follows. Section
\ref{sec:architecture} provides a brief architectural overview of GPUs. Section
\ref{sec:challenges} elaborates on the challenges specific to GPGPU
applications. Section \ref{sec:dynamic} discusses spawning granular helpers to
increase performance. Section \ref{sec:scheduling} deals with more efficient
scheduling within and among streaming multiprocessors, as well as tuning
hardware to match the current load of the GPU. Section \ref{sec:memory} talks
specifically about reducing memory contention. Section \ref{sec:software} is a
discussion of a few software and compiler solutions to the challenges of GPGPU.
Section
\ref{sec:conclusions} concludes the survey.

@@

Individual threads exist within a thread-block. These threads are grouped into
``warps'', usually around 32 at a time. The warp is the fundamental unit of
execution on GPUs. Threads within a warp are mapped to a vector functional unit
such that all threads execute the same instruction on different data, just a
like a SIMD unit. With this in mind, it would be more accurate to simply call
``threads'' a sequence of SIMD lane instructions, but this survey will continue
to refer to them as threads to maintain consistency with the literature. The
term ``Warp'' is the term used by NVIDIA, while AMD and OpenCL use
``wavefront''. Maxwell supports up to 64 warps resident per SM. Note that this
does not mean that all 64 warps are executing at the same time, but that the
architecture can support that many warp contexts concurrently. The SM maintains
a warp pool to store the context of all running threads. Warps are scheduled by
the warp schedulers from this pool.
